---
title: "Machine Learning - Analysis of weight-lifting exercises and building a predictive model for predicting the manner"
author: "Venkata Chaitanya Krishna Vasamsetty"
date: "07/26/2015"
output: html_document
---
# Data Processing
In order to ensure reproduceability, the following libraries are required and loaded. Also, e1071 package is additionally required while running this code.

```{r, echo=TRUE}
library(Hmisc)
library(foreach)
library(doParallel)
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
```
Also, the seed to set a specific value for reproduceability.
```{r, echo=TRUE}
set.seed(33333)
```
## Loading the Data
```{r, echo=TRUE}
trainingDataURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingDataURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```
The entries with "NA" and "#DIV/0!" are considered as null values while loading the datasets into memory.
```{r, echo=TRUE}
trainingData <- read.csv(url(trainingDataURL), na.strings=c("NA","#DIV/0!",""))
testingData <- read.csv(url(testingDataURL), na.strings=c("NA","#DIV/0!",""))
for(i in c(8:ncol(trainingData)-1)) {trainingData[,i] = as.numeric(as.character(trainingData[,i]))}
for(i in c(8:ncol(testingData)-1)) {testingData[,i] = as.numeric(as.character(testingData[,i]))}
```

Clean the variables that do not contribute well to prediction or had incomplete values in them.
```{r, echo=TRUE}
feature_set <- colnames(trainingData[colSums(is.na(trainingData)) == 0])[-(1:7)]
model_data <- trainingData[feature_set]
```
The training data is paritioned to two parts so that it can be later used for cross-validation.
```{r, echo=TRUE}
x <- createDataPartition(y=model_data$classe, p=0.6, list=FALSE)
PartitionedTrainingData <- model_data[x, ]; PartitionedTestingData <- model_data[-x, ]
```
## ML Algorithm for Prediction 

### Decision Tree
```{r, echo=TRUE}
modelFit_DecisionTree <- rpart(classe ~ ., data=PartitionedTrainingData, method="class")
fancyRpartPlot(modelFit_DecisionTree)
```
Confusion Matrix is used to test the prediction results
```{r, echo=TRUE}
predictions_DecisionTree <- predict(modelFit_DecisionTree, PartitionedTestingData, type = "class")
confusionMatrix(predictions_DecisionTree, PartitionedTestingData$classe) # an error might come up if e1071 is not installed. Please ensure that it is installed before running this line.
```

### Random Forests

```{r, echo=TRUE}
modelFit_RandomForests <- randomForest(classe ~. , data=PartitionedTrainingData)
```
Confusion Matrix is used to test the prediction results
```{r, echo=TRUE}
predictions_RandomForests <- predict(modelFit_RandomForests, PartitionedTestingData, type = "class")
confusionMatrix(predictions_RandomForests, PartitionedTestingData$classe)
```

## Conclusion
Random Forests provide better accuracy than Decision Trees.

## Generating answers for the test cases
This details on the approach used to find 
```{r, echo=TRUE}
predictions_testingData <- predict(modelFit_RandomForests, testingData, type = "class")
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictions_testingData)


```
